---
title: "Chapter 11 Practice"
author: "Trevor Yarnell"
date: today
format: html
editor: visual
---

## Task 0

```{r}
# Load required packages

library(dslabs)
library(dplyr)
```

## Task 1

```{r}
# Creates a new dataframe called C_Overload

C_Overload <- select(temp_carbon, year, carbon_emissions)

head(C_Overload)
```

Select only kept the year and carbon_emissions columns from the data-set. Head shows the first six rows of the new data frame.

## Task 2

```{r}
star_names <- stars %>%
  select(star)

head(star_names)
```

This chunk of code kept it as a data-frame.

## Task 3

```{r}
No_Florida <- murders %>%
  filter(state != "Florida")

head(No_Florida)
```

```{r}
unique(No_Florida$state)
```

By using unique you can check what states are included in the data-set and also ensure you're spelling is correct. When you run unique again I'm looking to see if Florida is still included.

## Task 4

```{r}
Fast_n_Furious <- mtcars %>%
  filter(am == 1, cyl == 8, qsec <= 15)

head(Fast_n_Furious)
nrow(Fast_n_Furious)
```

I used 1 which indicates that the car is a manual, 8 is the number of cylinders and \<= 15 means R picks out the cars that have a 1/4 times of less than or equal to 15.

## Task 5

```{r}
range(mtcars$disp)
```

```{r}
filtered_cars <- mtcars %>%
  filter(disp < 160 | disp > 250) %>%
  select(1:3)

all(filtered_cars$disp > 250 | filtered_cars$disp < 160)
```

## Task 7

select() allows you to pick out the columns you want and filter() allows you to filter our rows that match a condition you specify. \[\] also perform a similar function to both select() and filter().
