---
title: "Chapter 13 Notes"
subtitle: "group_by() and summarize()"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format:
  html:
    code-fold: false
---

# Loading Packages
```{r}
library(dslabs)
library(dplyr)
```

# Counts, Center, and Spread
Numeric variable: numbers that are measured or calculated
Categorical variable: stores labels or group names

Count: for a whole dataset means the total number of rows; grouped data mean the number of rows in each group

Center: typical value
Mean: average
Median: the middle value

Spread: how much the values vary
Standard deviation: how far values vary from the mean
Range: distance from the smallest to largest value

# count()
## Counting one varible
```{r}
mtcars %>%
  count(cyl)
```
Returns how many cars in the `mtcars` dataset have 4, 6, and 8 cylinders grouped by cylinder count.

## count() versus length()
length() counts how many values are in a vector
count() counts the rows in a df for each category or combination of categories

## count() with sorting
```{r}
mtcars %>%
  count(cyl, sort = TRUE)
```
Sorts the groups with the most vehicles to the least

## count() with multiple variables
```{r}
mtcars %>%
  count(cyl, am)
```

# summarize()
```{r}
mtcars %>%
  summarize(mean_mpg = mean(mpg))
```
Creates a new column that has the mean MPG for all the cars in the mtcars dataset

```{r}
mtcars %>%
  summarize(mean_mpg = mean(mpg),
            max_mpg = max(mpg),
            count = n())
```
## count() versus summarize()
```{r}
mtcars %>%
  group_by(cyl) %>%
  summarize(n = n())
```

# group_by()
```{r}
cyl_means <- mtcars %>%
  group_by(cyl) %>%
  summarize(mean_mpg = mean(mpg))
```

## `dplyr` default naming
```{r}
cyl_means <- mtcars %>%
  group_by(cyl) %>%
  summarize(mean(mpg))
```

# multiple statistics
```{r}
cyl_stats <- mtcars %>%
  group_by(cyl) %>%
  summarize(mean_mpg = mean(mpg),
            sd_mpg = sd(mpg),
            n = n())

cyl_stats
```

# multiple variables
```{r}
mtcars %>%
  group_by(cyl, am) %>%
  summarize(avg_mpg = mean(mpg),
            count = n())
```
## ungroup
```{r}
cyl_gear_stats <- mtcars %>%
  group_by(cyl, gear) %>%
  summarize(mean_mpg = mean(mpg))
```
```{r}
group_vars(cyl_gear_stats)
```
```{r}
is_grouped_df(cyl_gear_stats)
```
```{r}
cyl_gear_stats <- mtcars %>%
  group_by(cyl, gear) %>%
  summarize(mean_mpg = mean(mpg), .groups = "drop")

group_vars(cyl_gear_stats)
```
```{r}
is_grouped_df(cyl_gear_stats)
```
```{r}
cyl_gear_stats <- mtcars %>%
  group_by(cyl, gear) %>%
  summarize(mean_mpg = mean(mpg)) %>%
  ungroup()
```
```{r}
group_vars(cyl_gear_stats)
```
```{r}
is_grouped_df(cyl_gear_stats)
```


















