---
title: "Chapter 12 Notes"
subtitle: "mutate() and case_when()"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format:
  html:
    code-fold: true
execute:
  warning: true
  message: true
---

# Quarto Notes

# Load Packages
```{r}
library(dslabs)
library(dplyr)
```

# Vectorized Conditionals

## ifelse()
```{r}
x <- c(3, 2, 3, 6, 5)
```
```{r}
length(x)
```
`x` is a vector with the length of `{r} length(x)` rows.
```{r}
ifelse(test = x > 3, yes = "Large", no = "Small")
```
```{r}
length(ifelse(x > 3, "Large", "Small"))
```

## case_when()
```{r}
case_when(c(3, 2, 3, 6, 5) > 3 ~ "Large", .default = "Small")
```

Don't forget to load any required libraries when trying to use functions no in base R.

## More than Two Categories
```{r}
scores <- c(55, 65, 75, 85, 95)
```

### Nested ifelse()
```{r}
ifelse(scores < 60, "Low",
       ifelse(scores < 80, "Medium", "High"))
```

### Multiple Rule case_when()
```{r}
case_when(scores < 60 ~ "Low",
          scores < 80 ~ "Medium",
          TRUE ~ "High")
```
```{r}
case_when(scores < 80 ~ "Medium",
          scores < 60 ~ "Low",
          TRUE ~ "High")
```

# mutate()

## Column Copy
```{r}
names(mtcars)
```
```{r}
mtcars2 <- mtcars %>%
  mutate(new_column_copy = am)
```
```{r}
names(mtcars2)
```

### mutate() with Arithmetic
```{r}
measurements <- data.frame(height_in = c(60, 65, 70))
measurements
```

The `measurements` data frame has one column called `{r} colnames(measurements)`.

```{r}
measurements %>%
  mutate(height_cm = height_in * 2.54)
```

### case_when() inside mutate()
```{r}
student_scores <- data.frame(scores = c(55, 68, 72, 83, 91))
```
```{r}
student_scores <- student_scores %>%
  mutate(
    score_level = case_when(
      scores < 60 ~ "Low",
      scores < 80 ~ "Medium",
      TRUE ~ "High"
    )
  )
```
```{r}
head(student_scores)
```

`student_scores` is a fataframe with columns names: `{r} names (student_scores)`.

## Overwrite Column Values
```{r}
unique(mtcars$am)
```
```{r}
cars_overwrite <- mtcars %>%
  mutate(
    am = case_when(
      am == 0 ~ "Automatic",
      TRUE ~ "Manual"
    )
  )
```
```{r}
unique(cars_overwrite$am)
```

The `am` column in `cars_overwrite` has the unique values: `{r} unique(cars_overwrite$am)`.
