---
title: "Chapter 15 Practice"
subtitle: "Probability"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format:
  html:
    code-fold: false
---
# Set Seed
```{r}
set.seed(1)
```

# Task 1
Goal: Roll a 20 sided die 300 times and store in big_die
```{r}
big_die <- sample(1:20, size = 300, replace = TRUE)

any(big_die == 20)
```

# Task 2
## Step 1
Roll at d20 5221 times and store in `boring_weekend`.
```{r}
boring_weekend <- sample(1:20, 5221, TRUE)
```

## Step 2
Create a table that shows how many times each number was rolled and store it in `d20_counts`.
```{r}
d20_counts <- table(boring_weekend)

print(d20_counts)
```

## Step 3
Prportion that were 20.
```{r}
mean(boring_weekend == 20)

abs(mean(boring_weekend == 20) - 0.05)
```
About 5% of the rolls were nat 20s.

## Step 4
Simulate 100 newborn lengths, mean of 20 and standard deviation of 0.9 store as blanketed_babies.
```{r}
blanketed_babies <- rnorm(100, 20, 0.9)
```
```{r}
unique(blanketed_babies)
```

## Step 5
Check if any of babies are longer 23 inches and count how many.
```{r}
any(blanketed_babies > 23)
sum(blanketed_babies > 23)
```
Based on this simulation the hospital does not need longer blankets.

## Step 6
Create a vector called `weeks_too_long` that records how mnay babies are longer than 23 inches.
```{r}
weeks_two_long <- replicate(1000, sum(rnorm(100, 20, 0.9) > 23))
```
```{r}
table(weeks_two_long)
mean(weeks_two_long)
mean(weeks_two_long > 0)
max(weeks_two_long)
```
In a thousand weeks we only see 34 total weeks were babies are over 23 inches.

# Task 3
In step 6 of this practice section `replicate()` allowed me to repeatedly sample normal distributions. This enable me to run multiple iterations of by normal distribution looking for weeks where a baby is bigger than 23 inches. This is more efficent than manually rerunning the calculation a thousand times.








