---
title: "Chapter 15 Notes"
author: "Trevor Yarnell"
format: html
---

# Set Seed
`set.seed` helps to ensure reproducibility.
```{r}
set.seed(2025)
```

# Coin Flips
Goal is to flip a coin 1,000 times. 0 = heads and 1 = tails

Sample Space: A collection or set of all possible outcomes in a random experiment.

Set Notation: \(S\) = {Head, Tail}

```{r}
coins <- sample(x = 0:1, size = 1000, replace = TRUE)
```

Replacement is essential whe simulating independent outcomes, like coin flips.

There will be about 500 tails
```{r}
sum(coins)
```
There were `r sum(coins)` Tails in the coins vector.

# Dice Rolls
Goal is to simulate 10,000 roles of a fair 6-sides die and plot the results

```{r}
dice <- sample(1:6, 10000, replace = TRUE)
```

# Histograms

```{r}
hist(dice,
     breaks = seq(0.5, 6.5, by = 1),
     col = "lightblue",
     main = "10,000 Fair Die Rolls",
     xlab = "Face Value")
```
Updates to the code centers the bars on the x-axis rather than in bins.

Can also do bar plots, probably better for this type of discrete data.

```{r}
barplot(table(dice),
        main = "Counts by Die Face (Table Method)",
        col = "lightblue",
        xlab = "Face Value",
        ylab = "Count")
```

# Two Die

IF I roll two die I would expect to see a normal distribution centered around 7.

```{r}
sum_two <- sample(1:6, 10000, TRUE) + sample(1:6, 10000, TRUE)

hist(sum_two,
     breaks = seq(1.5, 12.5, by = 1),
     col = "magenta",
     main = "Sum of Two Dice (10,000 trials)",
     xlab = "Total")
```

# Symmetric Data

Simulating the heights of 1,000 adult females using `rnorm()`

```{r}
heights <- rnorm(n = 1000, mean = 64, sd = 2.5)
```

```{r}
hist(heights,
     breaks = seq(55, 75, by = 1),
     col = "green",
     main = "Simulated Female Heigths",
     xlab = "Height (inches)")
```

`rnow` will create slightly different data since it is sampling each time. The mean and sd may not match that of the parameters.

# Repeating Simulation

```{r}
tails_100 <- replicate(n = 100, expr = sum(sample(0:1, 5000, TRUE)))
```

Will flip a coin 5000 tiles then repeats that 100 times using `replicate()`.

```{r}
summary(tails_100)
```

```{r}
hist(tails_100,
     breaks = seq(2400, 2600, by = 25),
     col = "gold",
     main = "Distribution of Tails in 5,000 Coin Flips (100 reps)",
     xlab = "Count of Tails")
```

# Normal Data with `replicate()`

Generate 50 random numbers, compare each to zero and then count how many are below zero

```{r}
x <- rnorm(n = 50, mean = 2, sd = 3)
print(x)
```
```{r}
below_zero <- x < 0
print(below_zero)
```
```{r}
count <- sum(below_zero)
print(count)
```
```{r}
values_below_zero <- replicate(
  n = 200,
  expr = sum(rnorm(50, mean = 2, sd = 3) < 0)
)
```
```{r}
mean(values_below_zero)
```
```{r}
sd(values_below_zero)
```
```{r}
quantile(values_below_zero, c(.1, .5, .9))
```
```{r}
threshold <- 15
max_count <- max(values_below_zero)
breaks_int <- seq(2.5, max_count + 0.5, by = 1)

hist(values_below_zero,
     breaks = breaks_int,
     col = "pink",
     main = "Distribution of Opponents in 50-Student Surveys (200 reps)",
     xlab = "Count of Values Below Zero")

abline(v = threshold, col = "red", lwd = 2, lty = 2)
abline(v = mean(values_below_zero), col = "blue", lwd = 2, lty = 1)

legend("topright",
       legen = c("Halt Threshold (15)", "Mean Count"),
       col = c("red", "blue"),
       lty = c(2, 1),
       lwd = 2,
       bg = "white",
       bty = "o")
```

# Fair-Dice Check
Simulate rolling a dice with biased probabilities
```{r}
rolls <- sample(1:6, 1000, replace = TRUE, prob = c(.12, .13, .07, .23, .22, .23))
```

```{r}
round(prop.table(table(rolls)), 3)
```

## Fair vs. Unfair Dice
Simulate rolling two biased dice 1,000 times and compare to. two fair dice
```{r}
set.seed(8000)
biased <- c(.12, .13, .07, .23, .22, .23)
friend_sum <- sample(1:6, size = 1000, replace = TRUE, prob = biased) +
  sample(1:6, size = 1000, replace = TRUE, prob = biased)
```

```{r}
fair_sum <- sample(1:6, 1000, TRUE) + sample(1:6, 1000, TRUE)
```

```{r}
ploaded <- prop.table(table(friend_sum))
pfair <- prop.table(table(fair_sum))
```

```{r}
face_props <- rbind(Bias = ploaded, Fair = pfair)
barplot(face_props,
        beside = TRUE,
        col = c("red", "blue"),
        main = "Probabilities (2 Dice, 1000 rolls)",
        xlab = "Added Face Values",
        ylab = "Proportion")
legend("topright", legend = c("Loaded", "Fair"),
       fill = c("red", "blue"), bty = "n")
```


