---
title: "Chapter 12 Practice Space"
aauthor: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format:
  html:
    code-fold: true
execute:
  warning: true
  message: true
---

# Load Libraries
```{r}
library(dplyr)
library(dslabs)
```

# Concept Check
Which method returns a new vector with one result for every position?

b. Logical subsetting using `[ ]`

# Simple `ifelse()`
```{r}
?dslabs::heights
heights <- dslabs::heights
names(heights)
```

I am attempting to label the `height` in the `heights` dataset as "Tall" if it is greater than `68` and otherwise "Not Tall"

```{r}
height_label <- ifelse(
  test = heights$height > 68,
  yes = "Tall",
  no = "Not Tall"
)
```
```{r}
table(height_label)
```
There are `508` "Not Tall" and `542` "Tall".

# Recode with mutate() and case_when()

I am labeling the `mpg` of cars from the `mtcars` dataset as either "High MPG" if it is greater than `25`, "Low MPG" if it is less than `15`, and otherwise "Medium MPG".

```{r}
mpg_recode <- mtcars %>%
  mutate(mpg_class = case_when(
    mpg > 25 ~ "High MPG",
    mpg < 15 ~ "Low MPG",
    TRUE ~ "Medium MPG"
  ))
```
```{r}
table(mpg_recode$mpg_class)
```
There are `6` High MPG cars, `5` Low MPG cars, and `21` Medium MPG cars.

# Recode: Name Focus
```{r}
?dslabs::stars
stars <- dslabs::stars

colnames(stars)
summary(stars$temp)
```

I am recording the temperature variable in the stars dataset to indicate if the star is "Cool", Warm, or "Hot".

```{r}
stars_recode <- stars %>%
  mutate(temp_group = case_when(
    temp < 4000 ~ "Cool",
    temp <= 7500 ~ "Warm",
    TRUE ~ "Hot"
  ))
```
```{r}
table(stars_recode$temp_group)
```
There are `35` Cool stars, `38` Hot stars, and `23` Warm stars.

# Unit conversion with mutate()
```{r}
heights <- dslabs::heights
head(heights)
```

I am overwriting the height column to a more internationally friendly number.

```{r}
international_friendly_heights <- heights %>%
  mutate(height = height * 2.54)
```
```{r}
head(international_friendly_heights)
```
The values of the first six rows have been correctly converted to centimeters.












