---
title: "Chapter 13 Practice"
subtitle: "group_by() and summarize()"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format:
  html:
    code-fold: false
---

# Load Libraries
```{r}
library(dslabs)
library(dplyr)
```

# Single grouping
```{r}
data(stars)
?stars
colnames(stars)
```

I am calculating the `mean_temperature` and `mean_magnitude` from the `stars` dataset and storing it to `twinkle_twinkle`.
```{r}
twinkle_twinkle <- stars %>%
  group_by(type) %>%
  summarize(
    mean_temperature = mean(temp),
    mean_magnitude = mean(magnitude),
    .groups = "drop"
  )

twinkle_twinkle
```
The mean temperature and mean magnitude for stars in the dataset is 8762.556. The mean magnitude for stars in the dataset is 4.08...

# Single grouping + three stats
## Inspect the data
```{r}
data(olive)
?olive
colnames(olive)
```
I will be finding the mean, standard deviation of `oleic` fatty acid and counting the number of rows in each region.
```{r}
# fatty_acid <- "oleic"

Olive_Garden <- olive %>%
  group_by(region) %>%
  summarize(
    mean_value = mean(oleic),
    sd_value = sd(oleic),
    n = n()
  )

Olive_Garden
```
Code chunk does not operate when trying to store the type of fatty acid as a vector. Manually entered instead. Northern Italy has the largest `mean_value` for the fatty acid I selected.

# Counting unique values
```{r}
data("us_contagious_diseases")
?us_contagious_diseases

unique(us_contagious_diseases$disease)
length(unique(us_contagious_diseases$disease))
```
There are `r length(unique(us_contagious_diseases$disease))` unique diseases within the `disease` column of the `us_contagious_diseases` dataset. They are `r unique(us_contagious_diseases$disease)`

## Count the number of states in the dataset
```{r}
unique(us_contagious_diseases$state)
length(unique(us_contagious_diseases$state))
```
There are `r length(unique(us_contagious_diseases$state))` unique states within the `state` column of the `us_contagious_diseases` dataset.

# Make and test a prediction
Using group_by(state, disease) will return 357 groups. Each state will be broken into seven different groups for each disease.

I'm trying to test my theory above by creating a new df that groups the states and diseases.
```{r}
categorical_counts <- us_contagious_diseases %>%
  group_by(state, disease) %>%
  summarize()

count(categorical_counts)
```
My prediction was correct as there are 357 groups.

# National mean counts by disease
In this task I am calculating the total mean of the `count` for each `disease` and storing that in `Infectious_Burden`
```{r}
Infectious_Burden <- us_contagious_diseases %>%
  group_by(disease) %>%
  summarize(
    mean_count = mean(count, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
      arrange(desc(mean_count))

head(Infectious_Burden, 3)
```
Mumps and Measles have the highest national mean counts.

# Reflection
When using `group_by()` before `summarize()` it allows you to get descriptive statistics from different groups within the data. If you were to only use `summarize()` on a data set it would give you the mean, sd, etc for all rows and it would not be broken down by groups. For example earlier in this chapter I wrote code that calculated the. average MPG for all vehicle in the `mtcars` dataset shortly after that I create code that showed the average MPG by number of cylinders.












