---
title: "Signature Assignment"
subtitle: "Descriptive Statistics"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
execute: 
  warning: true
  message: true
---

# Part 1

## Load Packages and Data
```{r}
#| eval: false
install.packages("palmerpenguins")
```
```{r}
#| message: false
#| warning: false
library(dplyr)
library(palmerpenguins)
```

The contains information of three penguin species observed in the Palmer Archipelago, Antarctica, collected by Dr. Kristen Gorman with Palmer Station LTER

link: <https://allisonhorst.github.io/palmerpenguins/reference/penguins.html>

[Data Sources Reference](https://allisonhorst.github.io/palmerpenguins/reference/penguins.html)

```{r}
data("penguins")
```

## Data Types
```{r}
glimpse(penguins)
```
The `penguins` dataset contains `{r} nrow(penguins)` and `{r} ncol(penguins)` columns.

The `species` column is stored as `{r} class(penguins$species)` in R. The unique unordered values for `species` are: `{r} unique(penguins$species)`

The `island` column is stored as `{r} class(penguins$island)` in R. The unique unordered values for `island` are: `{r} unique(penguins$island)`

The `bill_length_mm` column is stored as `{r} class(penguins$bill_length_mm)` in R. 

The `bill_depth_mm` column is stored as `{r} class(penguins$bill_depth_mm)` in R. 

The `flipper_length_mm` column is stored as `{r} class(penguins$flipper_length_mm)` in R.

The `body_mass_g` column is stored as `{r} class(penguins$body_mass_g)` in R.

The `sex` column is stored as `{r} class(penguins$sex)` in R. The unique unordered values for `sex` are: `{r} unique(penguins$sex)`

The `year` column is stored as `{r} class(penguins$year)` in R. The unique ordered values for `year` are: `{r} unique(penguins$year)`

# Part 2

## Length

## Count 

## Mode

## Median

## Mean

# Part 3

# Spread

## Standard Deviation

## Range

## IQR

## Five Number Summary

# Part 4

## Describing Data with `dplyr`








