---
title: "Wildfire"
subtitle: "Wildfire Signature Assignment"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
---

# Load Data

1) Read `whp_raw_sample.csv` using readr, 2) Store in an object called `county_raw`, 3) Run `nrow` and `table` to verify data structure, 4) Explain what is in the data.

```{r}
library(readr) # Loads readr
library(dplyr)
county_raw <- read_csv("whp_raw_sample.csv") # Loads the dataset into an object
```
Had to consult AI to help me with this code chunk. Forgot to put the quotation marks around the file name which made it fail to load.

```{r}
glimpse(county_raw)
nrow(county_raw)
```
The dataset has four columns `whp`, `county`, `county_fips`, and `cell_area_km2`. It classifies what the wildfire risk is for a given area in a given county.

# Basic Data Audit Summaries

Create a summary of the data using `dplyr` summarize function and stoer that in a new object. Report the number of rows, number of missing values, minimum and maximum value of `cell_area_km2`.

```{r}
county_summary <- county_raw %>%
    summarize(
    n_rows = n(),
    n_na_whp = sum(is.na(whp)),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2)
  )
```

```{r}
county_summary
```

There are `r nrow(county_raw)` rows in the dataset, `r sum(is.na(county_raw$whp))` missing values,  and the minimum area is `r min(county_raw$cell_area_km2)` and the maximum area is `r max(county_raw$cell_area_km2)`. Each of the rows areas are close to each other.

# Compute WHP quantile edges

Use `quantile()` to compute the quantiles' of the `whp` column of the data and store in `e`.

```{r}
e <- quantile(
  county_raw$whp,
  probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1),
  na.rm = TRUE)
```

```{r}
e
```

The quantiles' show the distribution of the data for example the lowest 25% have a Wildfire Hazard Potential of 130.7305. This categorizes the data so that we can add labels to verbally communicate the risk.

To complete this task I needed to refer to previous notes.

# Add a categorical WHP risk label

Create a new data frame called `county_cats` by using `mutate()` and `case_when()` that categorizes the risk based on the percentile from Very Low to Ultra.

```{r}
county_cats <- county_raw %>%
  mutate(Category = case_when(
    whp >= e[1] & whp < e[2] ~ "Very Low",
    whp >= e[2] & whp < e[3] ~ "Low",
    whp >= e[3] & whp < e[4] ~ "Moderate",
    whp >= e[4] & whp < e[5] ~ "High",
    whp >= e[5] & whp < e[6] ~ "Very High",
    whp >= e[6] & whp < e[7] ~ "Extreme",
    whp >= e[7] & whp < e[8] ~ "Ultra (Top 1%)",
    TRUE ~ "NA"),
    Category = factor(
      Category,
      levels = c("Very Low", "Low", "Moderate", "High", "Very High", "Extreme", "Ultra (Top 1%)"),
      ordered = TRUE))
```
```{r}
county_cats %>% count(Category)
```
When using `case_when()` it compares the WHP value from the table and the one I've specified through edges. If the WHP values falls between the values I have set with the logical tests it specifies that column based on the categories I gave it.

# Aggregate area abd compute category probabilities

In this final task I am determining the aggregate fire risk by category using `group_by()`, `summarize()`, and `mutate()` and then calculating probability as a share of area. The result is then stored in `prob_tbl`.

```{r}
prob_tbl <- county_cats %>%
  group_by(Category) %>%
  summarize(area_km2 = sum(cell_area_km2),
            n_cells = n(),
            .groups = "drop") %>%
  mutate(
    Probability = area_km2/sum(area_km2)) %>%
      select(Category, Probability, area_km2, n_cells) # What is this actually doing? It's ordering how the columns in the table                                                              appear.
```
```{r}
prob_tbl
sum(prob_tbl$Probability)
```
This probabilty tables show that as the risk Category goes up so does the probability of a random ignition. It also shows that most of the area in Idaho is at Very Low to High risk with very little Very High and above.










