---
title: "Wildfire"
subtitle: "Wildfire Signature Assignment"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
---

# Load Data

1) Read `whp_raw_sample.csv` using readr, 2) Store in an object called `county_raw`, 3) Run `nrow` and `table` to verify data structure, 4) Explain what is in the data.

```{r}
library(readr) # Loads readr
county_raw <- read_csv("whp_raw_sample.csv") # Loads the dataset into an object
```
Had to consult AI to help me with this code chunk. Forgot to put the quotation marks around the code.

```{r}
head(county_raw) # Shows the first six rows of the dataset
summary(county_raw$whp) # Shows basic statistics of whp
```

The dataset has four columns `whp`, `county`, `county_fips`, and `cell_area_km2`. It classifies what the wildfire risk is for a given area in a given county.

# Basic Data Audit Summaries

Create a summary of the data using `dplyr` summarize function and stoer that in a new object. Report the number of rows, number of missing values, minimum and maximum value of `cell_area_km2`.

```{r}
library(dplyr) # Loads dplyr
county_summary <- county_raw %>%
  summarise(
    rows = nrow(county_raw),
    missing_values = sum(is.na(whp)),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2)
  )
```

```{r}
county_summary
```

There are `r nrow(county_raw)` rows in the dataset, `r sum(is.na(county_raw$whp))` missing values,  and the minimum area is `r min(county_raw$cell_area_km2)` and the maximum area is `r max(county_raw$cell_area_km2)`. Each of the rows areas are close to each other.

# Compute WHP quantile edges

Use `quantile()` to compute the quantiles' of the `whp` column of the data abd store in `e`.

```{r}
e <- quantile(county_raw$whp,
  probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1),
  na.rm = TRUE)
```

```{r}
e
```

The quantiles' show the distribution of the data for example the lowest 25% have a Wildfire Hazard Potential of 130.7305. This categorizes the data so that we can add labels to verbally communicate the risk.

To complete this task I needed to refer to previous notes.

# Add a categorical WHP risk label

I am going to create category labels for the different levels of Wildfire Risk Potential.

```{r}
county_cats <- county_raw %>%
  mutate(Category = case_when(
    is.na(whp) ~ "NA",
    whp >= 22046.7472 ~ "Ultra (Top 1%)",
    whp >= 7708.5383 ~ "Very High",
    whp >= 3479.8125 ~ "High",
    whp >= 1101.0156 ~ "Moderate",
    whp >= 397.6562 ~ "Low",
    whp >= 130.7305 ~ "Very Low"))
```

```{r}
head(county_cats)

```











