---
title: "Wildfire"
subtitle: "Wildfire Signature Assignment"
author: "Trevor Yarnell"
date: today
toc: true
toc-depth: 2
number-sections: true
format: html
---

# Part A

## Load Data

1) Read `whp_raw_sample.csv` using readr, 2) Store in an object called `county_raw`, 3) Run `nrow` and `table` to verify data structure, 4) Explain what is in the data.

```{r}
library(readr) # Loads readr
library(dplyr)
county_raw <- read_csv("whp_raw_sample.csv") # Loads the dataset into an object
```
Had to consult AI to help me with this code chunk. Forgot to put the quotation marks around the file name which made it fail to load.

```{r}
glimpse(county_raw)
nrow(county_raw)
```
The dataset has four columns `whp`, `county`, `county_fips`, and `cell_area_km2`. It classifies what the wildfire risk is for a given area in a given county.

## Basic Data Audit Summaries

Create a summary of the data using `dplyr` summarize function and stoer that in a new object. Report the number of rows, number of missing values, minimum and maximum value of `cell_area_km2`.

```{r}
county_summary <- county_raw %>%
    summarize(
    n_rows = n(),
    n_na_whp = sum(is.na(whp)),
    min_area = min(cell_area_km2),
    max_area = max(cell_area_km2)
  )
```

```{r}
county_summary
```

There are `r nrow(county_raw)` rows in the dataset, `r sum(is.na(county_raw$whp))` missing values,  and the minimum area is `r min(county_raw$cell_area_km2)` and the maximum area is `r max(county_raw$cell_area_km2)`. Each of the rows areas are close to each other.

## Compute WHP quantile edges

Use `quantile()` to compute the quantiles' of the `whp` column of the data and store in `e`.

```{r}
e <- quantile(
  county_raw$whp,
  probs = c(0, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1),
  na.rm = TRUE)
```

```{r}
e
```

The quantiles' show the distribution of the data for example the lowest 25% have a Wildfire Hazard Potential of 130.7305. This categorizes the data so that we can add labels to verbally communicate the risk.

To complete this task I needed to refer to previous notes.

## Add a categorical WHP risk label

Create a new data frame called `county_cats` by using `mutate()` and `case_when()` that categorizes the risk based on the percentile from Very Low to Ultra.

```{r}
county_cats <- county_raw %>%
  mutate(Category = case_when(
    whp >= e[1] & whp < e[2] ~ "Very Low",
    whp >= e[2] & whp < e[3] ~ "Low",
    whp >= e[3] & whp < e[4] ~ "Moderate",
    whp >= e[4] & whp < e[5] ~ "High",
    whp >= e[5] & whp < e[6] ~ "Very High",
    whp >= e[6] & whp < e[7] ~ "Extreme",
    whp >= e[7] & whp <= e[8] ~ "Ultra (Top 1%)",
    TRUE ~ "NA"),
    Category = factor(
      Category,
      levels = c("Very Low", "Low", "Moderate", "High", "Very High", "Extreme", "Ultra (Top 1%)"),
      ordered = TRUE))
```
```{r}
county_cats %>% count(Category)
```
When using `case_when()` it compares the WHP value from the table and the one I've specified through edges. If the WHP values falls between the values I have set with the logical tests it specifies that column based on the categories I gave it.

## Aggregate area abd compute category probabilities

In this final task I am determining the aggregate fire risk by category using `group_by()`, `summarize()`, and `mutate()` and then calculating probability as a share of area. The result is then stored in `prob_tbl`.

```{r}
prob_tbl <- county_cats %>%
  group_by(Category) %>%
  summarize(area_km2 = sum(cell_area_km2),
            n_cells = n(),
            .groups = "drop") %>%
  mutate(
    Probability = area_km2/sum(area_km2)) %>%
      select(Category, Probability, area_km2, n_cells) # What is this actually doing? It's ordering how the                                                                columns in the table appear.
```
```{r}
prob_tbl
sum(prob_tbl$Probability)
```
This probabilty tables show that as the risk Category goes up so does the probability of a random ignition. It also shows that most of the area in Idaho is at Very Low to High risk with very little Very High and above.

# Part B

## Tail probability with `plnorm()`

```{r}
p_high_exact <- plnorm(q = 10, meanlog = 1.907, sdlog = 0.280, lower.tail = FALSE)
```

```{r}
p_high_exact
```

## 95th Percentile Run Probability

```{r}
q95_high <- qlnorm(0.95, meanlog = 1.907, sdlog = 0.280)
```

```{r}
q95_high
```

## Calculating Density using `dlnorm()`

Use `dlnorm()` compute the density at 3km and 10km using the mean and SD from the High fire row on Table 1 and store in d_3 and d_10.

```{r}
d_3 <- dlnorm(3, meanlog = log(7), sdlog = log(2))
d_10 <- dlnorm(10, meanlog = log(7), sdlog = log(2))
```

```{r}
d_3
d_10
```

## Simulate High-hazard distances

```{r}
set.seed(2025)
High_Lognorm <- rlnorm(n = 450, meanlog = 1.907, sdlog = 0.280)
```

```{r}
summary(High_Lognorm)
length(High_Lognorm)
```
```{r}
fast_growing_count <- sum(High_Lognorm >= 10)
```

## Count fast-growing ignitions

```{r}
fast_growing_count
fast_growing_count/length(High_Lognorm)
```

## Visualize simulated distances

```{r}
hist(High_Lognorm,
     col = "Red",
     main = "Simulated 8-Hour Distances for High WHP (Log Normal Model)",
     xlab = "Distance")
```

# Part C

Research Question: In a season with 450 new ignitions statewide, what is the probability we see at least 50 fast-growing events (>= 10 km in 8 hours) when ignitions can occur in any hazard category according to their statewide area weights?

## Create category labels

Goal: Create a vector named `labels` from **Table 1** using `c()` and verify the labels.

```{r}
labels <- c("Very Low", "Low", "Moderate", "High", "Very High", "Extreme", "Ultra (Top 1%)")
```

```{r}
labels
```

Comment: I don't know why matching the text from the table will be helpful later, but I will soon.

## Create category probabilities

Goal: Store the probabilities from **Table 1** in a vector called `prob` and verify the probabilities round to one.

```{r}
prob <- c(.25, .25, .25, .15, .05, .04, .01)
```

```{r}
prob
sum(prob)
```

## Create lognormal parameter vectors

Goal: Store the Mean log value in a vector called `meanlog` and store the SD log values in a vector called `sdlog`.

```{r}
meanlog <- c(0.582, 1.046, 1.451, 1.907, 2.455, 2.866, 3.310)
sdlog <- c(0.472, 0.325, 0.325, 0.280, 0.246, 0.220, 0.212)
```

```{r}
meanlog
sdlog
length(meanlog) == length(sdlog)
```

Had to get some AI assistance with the `length(meanlog) == length(sdlog)`. Previously I only included one = sign and this gave me no output.

I see that as the mean goes up the sd goes down.

## Set global simulation constants

Goal: Set the constants provided from the research question in their respective vectors.

```{r}
N <- 450
S <- 100000
k <- 10
```

```{r}
N
S
k
```

## Sample categories for one season

Goal: Simulate one season of ignitions by using `sample()` with replacement and probability from `prob` and store in idx.

```{r}
idx <- sample.int(length(labels),
              size = N,
              replace = TRUE,
              prob
)
```

```{r}
head(idx)
table(labels[idx])
```

Comment: The lower the fire risk the more ignitions we see.

## Draw distances for one season

Goal: Use `rlnorm` to simulate the 8-hour distance

```{r}
d <- rlnorm(N, meanlog = meanlog[idx], sdlog = sdlog[idx])
```

```{r}
head(d)
summary(d)
```

Comment: The number of ignitions is much larger than the rest of the categories.

## Count fast runs for one season

Goal: Create a vector that marks which ignitions are fast-growing using `d >= k`. Then calculate the sum and compute the mean to get a fraction of fast ignitions.

```{r}
fast_flags_one <- d >= k
fast_count_one <- sum(fast_flags_one)
mean(fast_flags_one)
```

```{r}
fast_count_one
```

Comment: In this particular simulation I saw 56 fast runs.

## Bundle one-season logic in a single block

Goal: Bundle the code done in the previous task so that it just returns the total number of fast runs.

```{r}
fast_count_one <- {
  idx <- sample(length(labels),
                size = N,
                replace = TRUE,
                prob = prob)
  d <- rlnorm(N, meanlog = meanlog[idx],
              sdlog = sdlog[idx])
  sum(d >= k)
}
```

```{r}
fast_count_one
```

Comment: We can replicate this block many times to get different seasons.

## Use `replicate()` to simulate many seasons

Goal: Using `replicate()` simulate 5 seasons worth of fires

```{r}
set.seed(2025)
counts_fast_mix <- replicate(
  S,
  fast_count_one <- {
  idx <- sample(length(labels),
                size = N,
                replace = TRUE,
                prob = prob)
  d <- rlnorm(N, meanlog = meanlog[idx],
              sdlog = sdlog[idx])
  sum(d >= k)
}
)
```

```{r}
summary(counts_fast_mix)
mean(counts_fast_mix)
prob_at_least_50 <- mean(counts_fast_mix >= 50)
prob_at_least_50 * 100
```

Comment: The average number of fast run fires was about 45, the probability of a fast run was 22.563%.

## Visualize the distribution of seasonal fast counts

Goal: Create a histogram of `counts_fast_mix` and label the axes. Draw a line at the 50 mark.

```{r}
library(ggplot2) # Load ggplot2
```

```{r}
fast_count_df <- data.frame(fast_count = counts_fast_mix)
```

```{r}
fast_count_hist <- fast_count_df %>%
  ggplot(aes(x = fast_count)) +
  geom_histogram(aes(y = after_stat(density)),
                 binwidth = 1,
                 fill = "blue",
                 color = "white") +
  geom_vline(xintercept = 50,
             color = "red") +
  xlab("Number of Fast Ignitions") +
  ylab("Density")
```

```{r}
fast_count_hist
```

Comment: It supports that about only 22% of fires would be fast ignitions with the majority of seasons only having about 44 fast ignitions.










